---
title: "Practical Machine Learning"
author: "Nirpeksh Nandan"
date: "22 January 2019"
output: html_document
---


## R Markdown

The goal of the project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

The submission consists of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online.

You should also apply your machine learning algorithm to the 20 test cases available in the test data above. P

Reproducibility: Due to security concerns with the exchange of R code, your code will not be run during the evaluation by your classmates. Please be sure that if they download the repo, they will be able to view the compiled HTML version of your analysis.

Load libraries, download data, read into R and extract relevant information:

##Data Cleaning & Preprocessing
```{r }
library(caret)
library(ggplot2)
library(randomForest)

#download Data from the course provided links
train_url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(url=train_url, destfile="training.csv")

test_url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url=test_url, destfile="testing.csv")

#Import both training and testing datasets
train <- read.csv("training.csv", na.strings=c("NA","#DIV/0!",""))
test <- read.csv("testing.csv", na.strings=c("NA","#DIV/0!",""))

names(train)
str(train)
summary(train)
#The target prediction outcome
summary(train$classe)

inTrain <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
myTrain <- train[inTrain, ]
myTest <- train[-inTrain, ]
```
Afterwards, we need to clean and structure the data using only valid indicators/features. Acertain selection of feature will be removed. They are in the following criterias: 1.Variables with variances that are close to zero (considered constant) 2.Variables with missing readings over 75% 3.Variables that are not fit to be predictors
```{r}
mytrain_SUB <- myTrain
for (i in 1:length(myTrain)) {
  if (sum(is.na(myTrain[ , i])) / nrow(myTrain) >= .75) {
    for (j in 1:length(mytrain_SUB)) {
      if (length(grep(names(myTrain[i]), names(mytrain_SUB)[j]))==1) {
        mytrain_SUB <- mytrain_SUB[ , -j]
      }
    }
  }
}

dim(mytrain_SUB)

mytrain_SUB2 <- mytrain_SUB[,8:length(mytrain_SUB)]
NZV <- nearZeroVar(mytrain_SUB2, saveMetrics = TRUE)

keep <- names(mytrain_SUB2)
```
## Modelling & Prediction

```{r}
#Reproducibility set
set.seed(556)
#Random forest model fabrication
modFit <- randomForest(classe~., data = mytrain_SUB2)
print(modFit)
#cross validation on my testing data
#out of sample error
predict1 <- predict(modFit, myTest, type = "class")
confusionMatrix(myTest$classe, predict1)
#in sample error
predict_train <- predict(modFit, myTrain, type = "class")
confusionMatrix(myTrain$classe, predict_train)
```
##Summary
As shown in the accuracy result above, when tested on the subset data used for cross validation we get an accuracy of 99%. This can be used to estimate our out of sample error. On the other hand the fabricated model is fitted to the training portion of the data, we get an accuracy of 100%. This can be used to estimate the in sample error.
```{r}
predict_FINAL <- predict(modFit, test, type = "class")
print(predict_FINAL)
```